Extended Data

### 2022-03-23 doc/manuscript/Dow_manuscript_Nature.Rmd
update word count
#130

<!--LENGTH: The typical length of an article with 3-4 modest display items (figures and tables) is 2000-2500 words (summary paragraph plus body text). We're currently at 2672 words and 1 modest + 2 large display items. Articles typically have 30-50 references (currently at 50)-->

@@

In temperate deciduous forests, spring warming generally lengthens the period over which trees have photosynthetically active leaves[@jeong_phenology_2011;<!--@ibanez_forecasting_2010;@menzel_european_2006;@menzel_growing_1999;-->@lu_no_2022] and that over which the ecosystem is a net CO~2~ sink[@keenan_net_2014;@lu_no_2022].
Based on these observations, current terrestrial ecosystem models effectively assume that longer growing seasons lead to sustained increases in annual net CO~2~ uptake (i.e., net ecosystem exchange, NEE)[@buermann_widespread_2018;@zani_increased_2020;@keenan_net_2014;@ahlstrom_robustness_2012], as occurs when additional C is allocated to woody growth and therefore resides in the ecosystem for decades to centuries[@xue_global_2017;@russell_residence_2014].
However, model representation of C allocation to stem growth does not capture known decoupling of stem growth from photosynthate production[@delpierre_temperate_2016;@delpierre_wood_2016;@fatichi_moving_2014].
As a result, the consequences of rising spring temperatures on stem growth -- or woody aboveground net primary productivity (ANPP~woody~) on the ecosystem level -- may not be accurately represented in models[@delpierre_temperate_2016;@fatichi_moving_2014]. 
Understanding the sensitivity of ANPP~woody~ to spring temperatures is central to predicting the future of the temperate deciduous forest C sink.

@@

<!--conclusions-->
As spring temperatures become increasingly warmer, growing seasons will start earlier. 
However, barring rapid acclimation of temperate deciduous forests to the warming conditions, advancement in the timing of stem growth (Fig. 1) is unlikely to provide the sustained increase in CO~2~ sequestration in woody biomass and ensuant negative climate change feedback that is anticipated in most climate forecasting models[@buermann_widespread_2018;@zani_increased_2020;@keenan_net_2014;@ahlstrom_robustness_2012].
Rather, the dominant effect of rising temperatures on forest woody productivity will be a negative effect of high summer temperatures (Fig. `r tree_rings`), which constitutes a positive feedback to climate change.

>>>>

<!--LENGTH: The typical length of an article with 3-4 modest display items (figures and tables) is 2000-2500 words (summary paragraph plus body text). We're currently at 2652 words and 1 modest + 2 large display items. Articles typically have 30-50 references (currently at 50)-->

@@

In temperate deciduous forests, spring warming generally lengthens the period over which trees have photosynthetically active leaves[@jeong_phenology_2011;<!--@ibanez_forecasting_2010;@menzel_european_2006;@menzel_growing_1999;-->@lu_no_2022] and that over which the ecosystem is a net CO~2~ sink[@keenan_net_2014;@lu_no_2022].
Based on these observations, current terrestrial ecosystem models effectively assume that longer growing seasons contribute positively to annual net CO~2~ uptake (i.e., net ecosystem exchange, NEE)[@buermann_widespread_2018;@zani_increased_2020;@keenan_net_2014;@ahlstrom_robustness_2012], as occurs when additional C is allocated to woody growth and therefore resides in the ecosystem for decades to centuries[@xue_global_2017;@russell_residence_2014].
However, model representation of C allocation to stem growth does not capture known decoupling of stem growth from photosynthate production[@delpierre_temperate_2016;@delpierre_wood_2016;@fatichi_moving_2014].
As a result, the consequences of rising spring temperatures on stem growth -- or woody aboveground net primary productivity (ANPP~woody~) on the ecosystem level -- may not be accurately represented in models[@delpierre_temperate_2016;@fatichi_moving_2014]. 
Understanding the sensitivity of ANPP~woody~ to spring temperatures is central to predicting the future of the temperate deciduous forest C sink.

@@

<!--conclusions-->
As spring temperatures become increasingly warmer, growing seasons will start earlier. 
However, barring rapid acclimation of temperate deciduous forests to the warming conditions, advancement in the timing of stem growth (Fig. `r schematic_summary`) is unlikely to provide the sustained increase in CO~2~ sequestration in woody biomass and ensuant negative climate change feedback *that is anticipated in most climate forecasting models*[@buermann_widespread_2018;@zani_increased_2020;@keenan_net_2014;@ahlstrom_robustness_2012].
Rather, the dominant effect of rising temperatures on forest woody productivity will be a negative effect of high summer temperatures (Fig. `r tree_rings`), which constitutes a positive feedback to climate change.
